# Figures and Results

```{R include=FALSE}
options(rgl.useNULL = TRUE)

library(tidyverse)
library(evolvability)
library(evolqg)
library(here)
library(reshape2)
library(ggh4x)
library(ggrastr)
library(matrixcalc)

# utils (read matrices as lists, perform tests on each pair of matrices from lists)
source('./util/read_matrices.r')
source('./util/RSkewers.r')
source('./util/RDifferences.r')
source('./util/KrzCorsLast.r')
source('./util/KrzCorsFirst.r')
source('./util/MantelCors.r')
source('./util/AvgEvolvabilities.r')
source('./util/MaxEvolvabilities.r')
source('./util/Disparity.r')
source('./util/MantelTests.r')
source('./util/Cosines.r')

# return upper triangle of matrix (include diag)
uppertriang <- function(x) {
  x[upper.tri(x, diag = TRUE)]
}

# read in matrices csv
matrices_csv <- read.csv("small_families_matrices.csv")

# parameter combinations (adjust these to produce particular figures)
ns <- c(8) # options: 4, 8, 12
fs <- c(20, 60, 200) # options: large families = {50, 150, 500}, small families = {20, 60, 200}
hs <- c(0.1, 0.5) # options: 0.1, 0.5
multicolls <- c("low", "medium", "high") # options: low, medium, high
e_types <- c("ind", "sim", "opp") # options: ind, sim, opp

params <- expand.grid(n = ns, f = fs, h = hs, multicoll = multicolls, e_type = e_types)

adjust_p <- F # whether or not to adjust P matrices by heritability for relevant analyses (Ghat = h^2 .* P)
```

## Relative average evolvabilities
```{R}
# NOTE: change the h parameter in read_matrices() to scale P matrices by a different heritability value, e.g. h = 0.35

columns <- c("RelAvgEvolvability", "SE", "Ratio", "f", "Multicollinearity", "n", "h", "E")
evolvs <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(evolvs) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = adjust_p, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  evolv <- AvgEvolvabilities(a_list, g_list, p_list)
  
  evolvs <- rbind(
    evolvs,
    data.frame(
      RelAvgEvolvability = evolv[1, 1],
      SE = evolv[1, 2],
      Ratio = "P to true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      RelAvgEvolvability = evolv[2, 1],
      SE = evolv[2, 2],
      Ratio = "Estimated G to true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      RelAvgEvolvability = evolv[3, 1],
      SE = evolv[3, 2],
      Ratio = "Estimated G to P",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    )
  )
}

evolvs$E <- factor(evolvs$E, levels = c("ind", "sim", "opp"))
levels(evolvs$E) <- c("Independent", "Proportional", "Opposite")
evolvs$Heritability <- as.factor(evolvs$h)

dodge <- position_dodge(width = 0.15)

ggplot(data = evolvs) +
  geom_hline(aes(yintercept = 1), colour = "#000000", linetype="dashed") +
  geom_point(
    aes(
      x = as.factor(f),
      y = RelAvgEvolvability,
      color = Ratio,
      shape = Heritability
    ),
    position = dodge, cex = 2, alpha=0.9
  ) +
  geom_errorbar(
   aes(
      x = factor(f),
      y = RelAvgEvolvability,
      ymin = RelAvgEvolvability - (SE * 1.96),
      ymax = RelAvgEvolvability + (SE * 1.96),
      color = Ratio
    ),
    width = 0.2, position = dodge, alpha=0.9
  ) +
  labs(y = "Relative Average Evolvability", x = "Number of Families") +
  facet_nested(E ~ Multicollinearity + h, labeller = label_both) +
  scale_x_discrete(labels = c(50, 150, 500)) +
  theme_bw(base_size = 14) +
  theme(
    panel.spacing=unit(0, "lines"),
    panel.border=element_rect(color="grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  width = 10, height = 6,
  device = 'pdf', dpi = 500,
  filename = "avg_evolv_E_n8_adjusted.pdf"
)
```

## Relative maximum evolvabilities
```{R}
# NOTE: change the h parameter in read_matrices() to scale P matrices by a different heritability value, e.g. h = 0.35

columns <- c("RelMaxEvolvability", "SE", "Ratio", "f", "Multicollinearity", "n", "h", "E")
evolvs <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(evolvs) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = adjust_p, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  evolv <- MaxEvolvabilities(a_list, g_list, p_list)
  
  evolvs <- rbind(
    evolvs,
    data.frame(
      RelMaxEvolvability = evolv[1, 1],
      SE = evolv[1, 2],
      Ratio = "P to true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      RelMaxEvolvability = evolv[2, 1],
      SE = evolv[2, 2],
      Ratio = "Estimated G to true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      RelMaxEvolvability = evolv[3, 1],
      SE = evolv[3, 2],
      Ratio = "Estimated G to P",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    )
  )
}

evolvs$E <- factor(evolvs$E, levels = c("ind", "sim", "opp"))
levels(evolvs$E) <- c("Independent", "Proportional", "Opposite")
evolvs$Heritability <- as.factor(evolvs$h)

dodge <- position_dodge(width = 0.15)

ggplot(data = evolvs) +
  geom_hline(aes(yintercept = 1), colour = "#000000", linetype="dashed") +
  geom_point(
    aes(
      x = as.factor(f),
      y = RelMaxEvolvability,
      color = Ratio,
      shape = Heritability
    ),
    position = dodge, cex = 2, alpha=0.9
  ) +
  geom_errorbar(
   aes(
      x = factor(f),
      y = RelMaxEvolvability,
      ymin = RelMaxEvolvability - (SE * 1.96),
      ymax = RelMaxEvolvability + (SE * 1.96),
      color = Ratio
    ),
    width = 0.2, position = dodge, alpha=0.9
  ) +
  labs(y = "Relative Maximum Evolvability", x = "Number of Families") +
  facet_nested(n ~ Multicollinearity + h, labeller = label_both) +
  scale_x_discrete(labels = c(20, 60, 200)) +
  theme_bw(base_size = 14) +
  theme(
    panel.spacing=unit(0, "lines"),
    panel.border=element_rect(color="grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  width = 10, height = 6,
  device = 'pdf', dpi = 500,
  filename = "max_evolv_n_unadjusted.pdf"
)
```

## Krzanowski's Common Subspaces Similarities
```{R}
# n = 8 used for figures where E varies

columns <- c("Correlation", "SE", "Comparison", "f", "Multicollinearity", "n", "h", "E")
krzcors <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(krzcors) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = FALSE, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  krzcor <- KrzCorsFirst(a_list, g_list, p_list) # either KrzCorsFirst() or KrzCorsLast()
  
  krzcors <- rbind(
    krzcors,
    data.frame(
      Correlation = krzcor[1, 1],
      SE = krzcor[1, 2],
      Comparison = "P and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Correlation = krzcor[2, 1],
      SE = krzcor[2, 2],
      Comparison = "Estimated G and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Correlation = krzcor[3, 1],
      SE = krzcor[3, 2],
      Comparison = "Estimated G and P",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    )
  )
}

krzcors$E <- factor(krzcors$E, levels = c("ind", "sim", "opp"))
levels(krzcors$E) <- c("Independent", "Proportional", "Opposite")
krzcors$Heritability <- as.factor(krzcors$h)
krzcors$Multicollinearity <- factor(krzcors$Multicollinearity, levels = c("low", "medium", "high"))

dodge <- position_dodge(width = 0.15)

ggplot(data = krzcors) +
  geom_point(
    aes(
      x = as.factor(f),
      y = Correlation,
      color = Comparison,
      shape = Heritability
    ),
    position = dodge, cex = 2, alpha=0.9
  ) +
  geom_errorbar(
   aes(
      x = factor(f),
      y = Correlation,
      ymin = Correlation - (SE * 1.96),
      ymax = Correlation + (SE * 1.96),
      color = Comparison
    ),
    width = 0.2, position = dodge, alpha=0.9
  ) +
  labs(y = "Common Subspace Correlation", x = "Number of Families") +
  facet_nested(E ~ Multicollinearity + h, labeller = label_both) +
  ylim(0.0, 1.0) +
  scale_x_discrete(labels = c(50, 150, 500)) +
  theme_bw(base_size = 14) +
  theme(
    panel.spacing=unit(0, "lines"),
    panel.border=element_rect(color="grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  width = 10, height = 6,
  device = 'pdf', dpi = 500,
  filename = "krz_first_E_n8.pdf"
)
```

## Mantel tests (return proportion of p-values < 0.05)
```{R include=FALSE}
columns <- c("P", "Comparison", "Multicollinearity", "n", "f", "h", "E")
pvals <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(pvals) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = FALSE, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  pval <- MantelTests(a_list, g_list, p_list)
  pval <- round(pval, 3)
  
  pvals <- rbind(
    pvals,
    data.frame(
      P = pval[1, 1],
      Comparison = "P and true G",
      n = params[i, 1],
      f = params[i, 2],
      h = params[i, 3],
      Multicollinearity = params[i, 4],
      E = params[i, 5]
    ),
    data.frame(
      P = pval[2, 1],
      Comparison = "Estimated G and true G",
      n = params[i, 1],
      f = params[i, 2],
      h = params[i, 3],
      Multicollinearity = params[i, 4],
      E = params[i, 5]
    ),
    data.frame(
      P = pval[3, 1],
      Comparison = "Estimated G and P",
      n = params[i, 1],
      f = params[i, 2],
      h = params[i, 3],
      Multicollinearity = params[i, 4],
      E = params[i, 5]
    )
  )
}

pvals$E <- factor(pvals$E, levels = c("ind", "sim", "opp"))
levels(pvals$E) <- c("Independent", "Proportional", "Opposite")

write.csv(pvals, "mantel_tests_n4_E_large_fam.csv", row.names = FALSE)
```

## Mantel cors (Pearson matrix correlations)
```{R include=FALSE}
columns <- c("Correlation", "SE", "Comparison", "f", "Multicollinearity", "n", "h", "E")
mantelcors <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(mantelcors) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = FALSE, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  mantelcor <- MantelCors(a_list, g_list, p_list)
  
  mantelcors <- rbind(
    mantelcors,
    data.frame(
      Correlation = mantelcor[1, 1],
      SE = mantelcor[1, 2],
      Comparison = "P and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Correlation = mantelcor[2, 1],
      SE = mantelcor[2, 2],
      Comparison = "Estimated G and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Correlation = mantelcor[3, 1],
      SE = mantelcor[3, 2],
      Comparison = "Estimated G and P",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    )
  )
}

mantelcors$E <- factor(mantelcors$E, levels = c("ind", "sim", "opp"))
levels(mantelcors$E) <- c("Independent", "Proportional", "Opposite")
mantelcors$Heritability <- as.factor(mantelcors$h)
mantelcors$Multicollinearity <- factor(mantelcors$Multicollinearity, levels = c("low", "medium", "high"))

dodge <- position_dodge(width = 0.15)

ggplot(data = mantelcors) +
  geom_point(
    aes(
      x = as.factor(f),
      y = Correlation,
      color = Comparison,
      shape = Heritability
    ),
    position = dodge, cex = 2, alpha=0.9
  ) +
  geom_errorbar(
    aes(
      x = as.factor(f),
      y = Correlation,
      ymin = Correlation - (SE * 1.96),
      ymax = Correlation + (SE * 1.96),
      color = Comparison
    ),
    width = 0.2, position = dodge, alpha=0.9
  ) +
  labs(y = "Pearson Correlation", x = "Number of Families") +
  facet_nested(n ~ Multicollinearity + h, labeller = label_both) +
  scale_x_discrete(labels = c(50, 150, 500)) +
  theme_bw(base_size = 14) +
  theme(
    panel.spacing=unit(0, "lines"),
    panel.border=element_rect(color="grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  width = 10, height = 6,
  device = 'pdf', dpi = 500,
  filename = "mantel_cors_n4_E.pdf"
)
```

## Random Skewers
```{R}
columns <- c("Correlation", "SE", "Comparison", "Ne", "Multicollinearity", "n", "h", "E")
skewercors <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(skewercors) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = F, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  skewercor <- RSkewers(a_list, g_list, p_list)
  
  skewercors <- rbind(
    skewercors,
    data.frame(
      Correlation = skewercor[1, 1],
      SE = skewercor[1, 2],
      Comparison = "P and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Correlation = skewercor[2, 1],
      SE = skewercor[2, 2],
      Comparison = "Estimated G and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Correlation = skewercor[3, 1],
      SE = skewercor[3, 2],
      Comparison = "Estimated G and P",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    )
  )
}

skewercors$E <- factor(skewercors$E, levels = c("ind", "sim", "opp"))
levels(skewercors$E) <- c("Independent", "Proportional", "Opposite")
skewercors$Heritability <- as.factor(skewercors$h)
skewercors$Multicollinearity <- factor(skewercors$Multicollinearity, levels = c("low", "medium", "high"))

dodge <- position_dodge(width = 0.15)

ggplot(data = skewercors) +
  geom_point(
    aes(
      x = as.factor(f),
      y = Correlation,
      color = Comparison,
      shape = Heritability
    ),
    position = dodge, cex = 2, alpha=0.9
  ) +
  geom_errorbar(
    aes(
      x = as.factor(f),
      y = Correlation,
      ymin = Correlation - (SE * 1.96),
      ymax = Correlation + (SE * 1.96),
      color = Comparison
    ),
    width = 0.2, position = dodge, alpha=0.9
  ) +
  labs(y = "Random Skewer Correlation", x = "Number of Families") +
  facet_nested(E ~ Multicollinearity + h, labeller = label_both) +
  scale_x_discrete(labels = c(50, 150, 500)) +
  theme_bw(base_size = 14) +
  theme(
    panel.spacing=unit(0, "lines"),
    panel.border=element_rect(color="grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  width = 10, height = 6,
  device = 'pdf', dpi = 500,
  filename = "random_skewers_E_n4.pdf"
)
```

## Response Differences
```{R}
columns <- c("Difference", "SE", "Comparison", "f", "Multicollinearity", "n", "h", "E")
rdiffs <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(rdiffs) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params[i, 1],
      f == params[i, 2],
      multicoll == params[i, 4],
      h == params[i, 3],
      e_type == params[i, 5]
    ) %>%
    read_matrices(scale_p = adjust_p, h = params[i, 3])
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  rd <- RDifferences(a_list, g_list, p_list)
  
  rdiffs <- rbind(
    rdiffs,
    data.frame(
      Difference = 100*rd[1, 1],
      SE = rd[1, 2],
      Comparison = "P and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Difference = rd[2, 1],
      SE = rd[2, 2],
      Comparison = "Estimated G and true G",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    ),
    data.frame(
      Difference = rd[3, 1],
      SE = rd[3, 2],
      Comparison = "Estimated G and P",
      n = params[i, 1],
      f = params[i, 2],
      Multicollinearity = params[i, 4],
      h = params[i, 3],
      E = params[i, 5]
    )
  )
}

rdiffs$E <- factor(rdiffs$E, levels = c("ind", "sim", "opp"))
levels(rdiffs$E) <- c("Independent", "Proportional", "Opposite")
rdiffs$Heritability <- as.factor(rdiffs$h)
rdiffs$Multicollinearity <- factor(rdiffs$Multicollinearity, levels = c("low", "medium", "high"))

dodge <- position_dodge(width = 0.15)

ggplot(data = rdiffs) +
  geom_point(
    aes(
      x = as.factor(f),
      y = Difference,
      color = Comparison,
      shape = Heritability
    ),
    position = dodge, cex = 2, alpha=0.9
  ) +
  geom_errorbar(
    aes(
      x = as.factor(f),
      y = Difference,
      ymin = Difference - (SE * 1.96),
      ymax = Difference + (SE * 1.96),
      color = Comparison
    ),
    width = 0.2, position = dodge, alpha=0.9
  ) +
  labs(y = "Proportional Response Difference (%)", x = "Number of Families") +
  facet_nested(E ~ Multicollinearity + h, labeller = label_both) +
  scale_x_discrete(labels = c(50, 150, 500)) +
  theme_bw(base_size = 14) +
  theme(
    panel.spacing=unit(0, "lines"),
    panel.border=element_rect(color="grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  width = 10, height = 6,
  device = 'pdf', dpi = 500,
  filename = "rdiffs_E_n4.pdf"
)
```

## Disparity
```{r}
columns <- c("Disparity", "SE", "Comparison", "# of Traits", "# of Families", "Heritability", "Multicollinearity", "E Type")
disparities <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(disparities) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params$n[i],
      f == params$f[i],
      multicoll == params$multicoll[i],
      h == params$h[i],
      e_type == params$e_type[i]
    ) %>%
    read_matrices(scale_p = FALSE, h = params$h[i], cor = TRUE)
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  disp <- Disparity(a_list, g_list, p_list)
  disp <- round(disp, 4)
  
  disparities <- rbind(
    disparities,
    data.frame(
      Disparity = disp[1, 1],
      SE = disp[1, 2],
      Comparison = "P and true G",
      `# of Traits` = params$n[i],
      `# of Families` = params$f[i],
      Heritability = params$h[i],
      Multicollinearity = params$multicoll[i],
      `E Type` = params$e_type[i]
    ),
    data.frame(
      Disparity = disp[2, 1],
      SE = disp[2, 2],
      Comparison = "Estimated G and true G",
      `# of Traits` = params$n[i],
      `# of Families` = params$f[i],
      Heritability = params$h[i],
      Multicollinearity = params$multicoll[i],
      `E Type` = params$e_type[i]
    ),
    data.frame(
      Disparity = disp[3, 1],
      SE = disp[3, 2],
      Comparison = "P and estimated G",
      `# of Traits` = params$n[i],
      `# of Families` = params$f[i],
      Heritability = params$h[i],
      Multicollinearity = params$multicoll[i],
      `E Type` = params$e_type[i]
    )
  )
}

write.csv(disparities, "disparity_results_n8_E_small_fam.csv", row.names = FALSE)
```

## Cosine similarities between leading eigenvectors
```{r}
columns <- c("Cosine", "SE", "Comparison", "# of Traits", "# of Families", "Heritability", "Multicollinearity", "E Type")
cosines <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(cosines) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params$n[i],
      f == params$f[i],
      multicoll == params$multicoll[i],
      h == params$h[i],
      e_type == params$e_type[i]
    ) %>%
    read_matrices(scale_p = FALSE, h = params$h[i], cor = F)
  
  a_list <- matrices[[1]]
  g_list <- matrices[[2]]
  p_list <- matrices[[3]]
  
  cos <- Cosines(a_list, g_list, p_list)
  cos <- round(cos, 2)
  
  cosines <- rbind(
    cosines,
    data.frame(
      Cosine = cos[1, 1],
      SE = cos[1, 2],
      Comparison = "P and true G",
      `# of Traits` = params$n[i],
      `# of Families` = params$f[i],
      Heritability = params$h[i],
      Multicollinearity = params$multicoll[i],
      `E Type` = params$e_type[i]
    ),
    data.frame(
      Cosine = cos[2, 1],
      SE = cos[2, 2],
      Comparison = "Estimated G and true G",
      `# of Traits` = params$n[i],
      `# of Families` = params$f[i],
      Heritability = params$h[i],
      Multicollinearity = params$multicoll[i],
      `E Type` = params$e_type[i]
    ),
    data.frame(
      Cosine = cos[3, 1],
      SE = cos[3, 2],
      Comparison = "P and estimated G",
      `# of Traits` = params$n[i],
      `# of Families` = params$f[i],
      Heritability = params$h[i],
      Multicollinearity = params$multicoll[i],
      `E Type` = params$e_type[i]
    )
  )
}

write.csv(cosines, "cosine_results_n4_E_large_fam.csv", row.names = FALSE)
```

## Regression (Plots)
```{r}
columns <- c("Estimated Genetic Correlation", "True Genetic Correlation", "Comparison", "n", "f", "h", "multicoll", "E")
comparison_df <- data.frame(matrix(ncol = length(columns), nrow = 0))
colnames(comparison_df) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params$n[i],
      f == 200,
      multicoll == "medium",
      h == params$h[i],
      e_type == params$e_type[i]
    ) 
  
  if (nrow(matrices)==0) {
    next
  }
  
  matrices <- matrices %>%
    read_matrices(scale_p = FALSE, cor = TRUE)
  
  a_matrices <- matrices[[1]]
  g_matrices <- matrices[[2]]
  p_matrices <- matrices[[3]]
  
  for (j in seq_along(a_matrices)) {
    a_upper <- a_matrices[[j]][upper.tri(a_matrices[[j]])]
    g_upper <- g_matrices[[j]][upper.tri(g_matrices[[j]])]
    p_upper <- p_matrices[[j]][upper.tri(p_matrices[[j]])]
    
    a_upper_melted <- melt(a_upper)
    g_upper_melted <- melt(g_upper)
    p_upper_melted <- melt(p_upper)
    
    comparison_data_g <- data.frame(
      "Estimated Genetic Correlation" = g_upper_melted$value,
      "True Genetic Correlation" = a_upper_melted$value,
      "Comparison" = "Estimated G and true G",
      "n" = params$n[i],
      "f" = params$f[i],
      "h" = params$h[i],
      "multicoll" = params$multicoll[i],
      "E" = params$e_type[i]
    )
    
    comparison_df <- rbind(comparison_df, comparison_data_g)
    
    comparison_data_p <- data.frame(
      "Estimated Genetic Correlation" = p_upper_melted$value,
      "True Genetic Correlation" = a_upper_melted$value,
      "Comparison" = "P and true G",
      "n" = params$n[i],
      "f" = params$f[i],
      "h" = params$h[i],
      "multicoll" = params$multicoll[i],
      "E" = params$e_type[i]
    )
    
    comparison_df <- rbind(comparison_df, comparison_data_p)
  }
}

comparison_df$E <- factor(comparison_df$E, levels = c("ind", "sim", "opp"))
levels(comparison_df$E) <- c("Independent", "Proportional", "Opposite")
comparison_df$h <- factor(comparison_df$h, levels = c(0.1, 0.5))

points_df <- comparison_df %>%
  group_by(E, n, h, Comparison) %>%
  slice_sample(prop = 1) %>%          
  slice_head(n = 2000) %>%     
  ungroup()

ggplot(comparison_df, aes(x = True.Genetic.Correlation,
                          y = Estimated.Genetic.Correlation)) +
  geom_point(data = points_df, aes(color = Comparison),
             size = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  coord_cartesian(ylim = c(-1, 1)) +
  facet_nested(E + n ~ h + Comparison, labeller = label_both) +
  theme_bw(base_size = 11) +
  labs(y = "Estimated G or P Correlations", x = "True G Correlations") +
  guides(color = "none") +
  theme(
    panel.spacing = unit(0, "lines"),
    panel.border = element_rect(color = "grey10"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(
  filename = "regression_plots_all.pdf",
  width = 10, height = 8,
  device = 'pdf'
)
```

## Regression (Tables)
```{R}
columns <- c("Estimated Genetic Correlation", "True Genetic Correlation", "Comparison", "n", "f", "h", "multicoll", "E")
comparison_df <- data.frame(matrix(ncol = length(columns), nrow = 0))
colnames(comparison_df) <- columns

for (i in seq_len(nrow(params))) {
  matrices <- matrices_csv %>%
    filter(
      n == params$n[i],
      f == 200,
      multicoll == "medium",
      h == params$h[i],
      e_type == params$e_type[i]
    ) %>%
    read_matrices(scale_p = FALSE, cor = TRUE)
  
  a_matrices <- matrices[[1]]
  g_matrices <- matrices[[2]]
  p_matrices <- matrices[[3]]
  
  for (j in seq_along(a_matrices)) {
    a_upper <- a_matrices[[j]][upper.tri(a_matrices[[j]])]
    g_upper <- g_matrices[[j]][upper.tri(g_matrices[[j]])]
    p_upper <- p_matrices[[j]][upper.tri(p_matrices[[j]])]
    
    comparison_data_g <- data.frame(
      "Estimated Genetic Correlation" = g_upper,
      "True Genetic Correlation" = a_upper,
      "Comparison" = "Estimated G and true G",
      "n" = params$n[i],
      "f" = params$f[i],
      "h" = params$h[i],
      "multicoll" = params$multicoll[i],
      "E" = params$e_type[i]
    )
    
    comparison_data_p <- data.frame(
      "Estimated Genetic Correlation" = p_upper,
      "True Genetic Correlation" = a_upper,
      "Comparison" = "P and true G",
      "n" = params$n[i],
      "f" = params$f[i],
      "h" = params$h[i],
      "multicoll" = params$multicoll[i],
      "E" = params$e_type[i]
    )
    
    comparison_df <- rbind(comparison_df, comparison_data_g, comparison_data_p)
  }
}

regression_results <- comparison_df %>%
  group_by(n, f, h, multicoll, E, Comparison) %>%
  do({
    model <- lm(Estimated.Genetic.Correlation ~ True.Genetic.Correlation, data = .)
    print(summary(model))
    data.frame(
      intercept = coef(model)[1],
      slope = coef(model)[2],
      p_value = summary(model)$coefficients[2, 4],
      r_squared = summary(model)$r.squared
    )
  }) %>%
  ungroup()

write.csv(regression_results, "figS1_regression_results_small_fam2.csv", row.names = FALSE)
```

## Error proportion
```{r}
# build dataframe
columns <- c("Ratio", "SE", "Comparison", "# of Traits", "# of Families", "Heritability", "Multicollinearity", "E Type")
ratiodf <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(ratiodf) <- columns

for (i in 1:nrow(params)) {
  matrices <- matrices_csv %>%
    filter(
      n == params$n[i],
      f == params$f[i],
      multicoll == params$multicoll[i],
      h == params$h[i],
      e_type == params$e_type[i]
    ) %>%
    read_matrices(scale_p = adjust_p, h = params$h[i])
  
  a_matrices <- matrices[[1]]
  g_matrices <- matrices[[2]]
  p_matrices <- matrices[[3]]
  
  # norm of differences between Ghat and true G matrices (Frobenius norm)
  map2_dbl(p_matrices, a_matrices, ~frobenius.norm(.x - .y)) -> p_diff
  map2_dbl(g_matrices, a_matrices, ~frobenius.norm(.x - .y)) -> g_diff

  ratios <- round(g_diff / p_diff, 2)
  mean_ratio <- mean(ratios)
  se_ratio <- sd(ratios) / sqrt(length(ratios))
  
  # add to df
  ratiodf <- rbind(ratiodf, data.frame(Ratio = mean_ratio, SE = se_ratio, Comparison = "F_G/F_P", `# of Traits` = params$n[i], `# of Families` = params$f[i], Heritability = params$h[i], Multicollinearity = params$multicoll[i], `E Type`=params$e_type[i]))
}

write.csv(ratiodf, "norm_small_fam.csv", row.names = FALSE)
```

## Response Differences (Adjustments by estimated heritability)
```{R}
# RUN THIS CHUNK LAST

ns <- c(4)
fs <- c(500)
hs <- c(0.5)
multicolls <- c("low", "medium", "high")
e_types <- c("opp", "ind", "sim")

# all combinations of parameters
params <- expand.grid(ns, fs, hs, multicolls, e_types)

# build dataframe
columns <- c("Difference", "SE", "Comparison", "Ne", "Multicollinearity", "n", "h", "E")
rdiffs <- data.frame(matrix(nrow = 0, ncol = length(columns)))
colnames(rdiffs) <- columns

for (i in 1:nrow(params)) {
  # read matrices
  matrices1 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = F)
  matrices2 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.1)
  matrices3 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.2)
  matrices4 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.3)
  matrices5 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.4)
  matrices6 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.5)
  matrices7 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.6)
  matrices8 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.7)
  matrices9 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.8)
  matrices10 <- read.csv("large_families_matrices.csv") %>% filter(n==params[i, 1]&f==params[i, 2]&multicoll==params[i,4]&h==params[i, 3]&e_type==params[i, 5]) %>% read_matrices(scale_p = T, h=0.9)
  
  a_list <- matrices1[[1]]
  g_list <- matrices1[[2]]
  p_list1 <- matrices1[[3]]
  p_list2 <- matrices2[[3]]
  p_list3 <- matrices3[[3]]
  p_list4 <- matrices4[[3]]
  p_list5 <- matrices5[[3]]
  p_list6 <- matrices6[[3]]
  p_list7 <- matrices7[[3]]
  p_list8 <- matrices8[[3]]
  p_list9 <- matrices9[[3]]
  p_list10 <- matrices10[[3]]
  
  # get response differences
  rd1 <- RDifferences(a_list, g_list, p_list1)
  rd2 <- RDifferences(a_list, g_list, p_list2)
  rd3 <- RDifferences(a_list, g_list, p_list3)
  rd4 <- RDifferences(a_list, g_list, p_list4)
  rd5 <- RDifferences(a_list, g_list, p_list5)
  rd6 <- RDifferences(a_list, g_list, p_list6)
  rd7 <- RDifferences(a_list, g_list, p_list7)
  rd8 <- RDifferences(a_list, g_list, p_list8)
  rd9 <- RDifferences(a_list, g_list, p_list9)
  rd10 <- RDifferences(a_list, g_list, p_list10)
  
  # add to df
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd1[1,1], SE = 100*rd1[1,2], Comparison = "P (not adjusted) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=1.0))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd2[1,1], SE = 100*rd2[1,2], Comparison = "P (adjusted, error=0.4) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.1))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd3[1,1], SE = 100*rd3[1,2], Comparison = "P (adjusted, error=0.3) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.2))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd4[1,1], SE = 100*rd4[1,2], Comparison = "P (adjusted, error=0.2) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.3))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd5[1,1], SE = 100*rd5[1,2], Comparison = "P (adjusted, error=0.1) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.4))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd6[1,1], SE = 100*rd6[1,2], Comparison = "P (adjusted, error=0.0) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.5))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd7[1,1], SE = 100*rd6[1,2], Comparison = "P (adjusted, error=-0.1) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.6))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd8[1,1], SE = 100*rd6[1,2], Comparison = "P (adjusted, error=-0.2) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.7))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd9[1,1], SE = 100*rd6[1,2], Comparison = "P (adjusted, error=-0.3) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.8))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd10[1,1], SE = 100*rd6[1,2], Comparison = "P (adjusted, error=-0.4) and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0.9))
  rdiffs <- rbind(rdiffs, data.frame(Difference = 100*rd10[2,1], SE = 100*rd6[2,2], Comparison = "Ghat and true G", n=params[i, 1], Ne=params[i, 2]*params[i, 3], Multicollinearity=params[i, 4], h=params[i, 3], E=params[i, 5], h_guess=0))
}

rdiffs$E <- factor(rdiffs$E, levels = c("ind", "sim", "opp"))
levels(rdiffs$E) <- c("Independent", "Proportional", "Opposite")

rdiffs$Heritability <- as.factor(rdiffs$h)
rdiffs$Multicollinearity <- factor(rdiffs$Multicollinearity, levels = c("low", "medium", "high"))
rdiffs$Comparison <- factor(
  rdiffs$Comparison,
  levels = c(
    "P (adjusted, error=0.4) and true G",
    "P (adjusted, error=0.3) and true G",
    "P (adjusted, error=0.2) and true G",
    "P (adjusted, error=0.1) and true G",
    "P (adjusted, error=0.0) and true G",
    "P (adjusted, error=-0.1) and true G",
    "P (adjusted, error=-0.2) and true G",
    "P (adjusted, error=-0.3) and true G",
    "P (adjusted, error=-0.4) and true G",
    "P (not adjusted) and true G"
  )
)

# plot data
ggplot(data=rdiffs) +
  geom_point(aes(x=factor(h_guess), y=Difference), cex=2) +
  geom_errorbar(aes(x=factor(h_guess), y=Difference, ymin=Difference-(SE*1.96), ymax=Difference+(SE*1.96)), width=.1) +
  labs(y = "Proportional Response Difference (%)", x = "Heritability Estimate Error (%)") +
   scale_x_discrete(
    limits = c("0.1", "0.2", "0.3", "0.4", "0.5", "0.6", "0.7", "0.8", "0.9", "1"),
    labels = c("-80", "-60", "-40", "-20", "0", "20", "40", "60", "80", "100")
  ) +
  geom_hline(
    data = subset(rdiffs, h_guess == 0),
    aes(yintercept = Difference),
    linetype = "dashed",
    colour   = "black",
    linewidth = 1.0
  ) +
  geom_rect(
    data = subset(rdiffs, h_guess == 0),
    aes(
      xmin = -Inf, xmax = Inf,
      ymin = Difference - 1.96 * SE,
      ymax = Difference + 1.96 * SE
    ),
    fill = "black", alpha = 0.1,
    inherit.aes = FALSE
  ) +
  facet_grid(E~Multicollinearity, labeller = label_both) +
  theme_bw(base_size=14) +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "none")

ggsave(width = 10, height = 8, device='pdf', dpi=500, filename = "rdiff_adjustments.pdf")
```